---
title: "Analysis of Google Play Store Apps"
author: "Minhaz Khan"
date: "January 12, 2019"
output: html_document
---

```{r}
library(tidyverse)
```


##Reading in the Data
```{r}
apps = read_csv("googleplaystore.csv")
head(apps)
```

##Average rating of each genre

Filtered out the missing rating values and stored them in a new data frame which we will use for the rest of the analysis.

```{r}
apps %>% filter(Rating != 'NaN') -> true_apps
true_apps %>% group_by(Genres) %>% summarise(avg=mean(Rating))
```
Getting a glimpse of the data to convert the numerical data from being characters
```{r}
glimpse(true_apps)
```

##Seperating Numerical and Character values from columns
```{r}
true_apps %>% filter(Size != "Varies with device") %>% 
  separate(Size, c("Size","Type"), sep = -1, convert = TRUE) %>% 
  separate(Installs, c("Installs","Symbol"), sep = -1, convert = TRUE) -> apps2

apps2$Symbol = NULL
apps2$Category = NULL
apps2$`Current Ver`= NULL
```
Removed Symbol because it was just character; removed category because it was the same as genre, removed Current Version because it isn't as necessary as when the app was last updated. Kept Android version because exploring app compatibility with OS might be interesting.

##Converting character values to numeric
```{r}
apps2$Installs = as.numeric(gsub(",","",apps2$Installs))
apps2$Size = as.numeric(as.character(apps2$Size))
glimpse(apps2)
```
Converting dates
```{r}
library(lubridate)
apps2$`Last Updated` = gsub(",","",apps2$`Last Updated`)
apps2 %>% mutate(`Last Update` = mdy(`Last Updated`))
View(apps2)
```

